### 2021-04-02 paper/glossary.tex
Finished Introduction and Aims and Obj

%%%%%%%%%%%%%%%%%%
% TERMS
%%%%%%%%%%%%%%%%%%
\newglossaryentry{stub}
{
        name 			= stub,
        description	= {A piece of code that is used to marshal parameters for transmission across the network}
}
\newglossaryentry{phishing}
{
        name 			= phishing,
        description	= {Sending an email that falsely claims to be from a legitimate organisation, usually combined with a threat or request for information}
}
\newglossaryentry{smishing}
{
        name			= smishing,
        description 	= {Sending a text message via SMS that falsely claims to be from a legitimate organisation, usually containing a link to a malicious website}
}
\newglossaryentry{malware}
{
        name			= malware,
        description	= {Malicious computer software that interferes with normal computer function or sends personal data about the user to unauthorised parties}
}
\newglossaryentry{cryptocurrency}
{
        name			= cryptocurrency,
        description	= {A digital currency produces by a public network}
}
\newglossaryentry{blockchain}
{
        name			= blockchain,
        description	= {A growing list of records, called blocks, that are linked using cryptography}
}
\newglossaryentry{ledger}
{
        name			= ledger,
        description	= {A record of all transactions executed on a particular cryptocurrency}
}
\newglossaryentry{middleware}
{
        name			= middleware,
        description	= {Software that functions at an intermediate layer between applications and the operating system to provide distributed functions}
}
\newglossaryentry{hashing}
{
        name			= hashing,
        description	= {The practise of taking data and representing that data as a fixed-length string}
}
%%%%%%%%%%%%%%%%%%
% ACRONYMS
%%%%%%%%%%%%%%%%%%
\newacronym{APT}{APT}{Advanced Persistent Threat}
\newacronym{NCSC}{NCSC}{National Cyber Security Center}
\newacronym{OS}{OS}{Operating System}